---
title: "Text analysis of Robert Louis Stevenson's Treasure Island"
description: |
  Project completed for an assignment in Allison Horst's ESM 244 class.
author:
  - name: Quin Smith
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
library(gutenbergr)
library(jpeg)
library(here)
library(ggpubr)
```

### Introduction

The following text analysis-- with associated Rmarkdown file-- explores word frequency trends and conveys the AFINN sentiment score of each chapter of Robert Louis Stevenson's classic pirate adventure novel, *Treasure Island.*   

```{r, warning = FALSE, message = FALSE}
#read in Treasure Island by Robert Louis Stevenson

ti <- gutenberg_download(120, mirror = "http://mirrors.xmission.com/gutenberg/")
```

```{r, warning = FALSE, message=FALSE}
#trim the text/tidy
ti_tidy <- ti %>% 
  mutate(text = str_trim(text))
```

```{r, warning = FALSE, message = FALSE}
#remove text outside of chapters in the novel (remove the Part structure), turn chapters into numerics 
ti_df <- ti_tidy %>% 
  slice(-(1:110)) %>% 
  mutate(chapter = case_when(
    str_detect(text, "^[0-9]+$") ~ text,
    str_detect(text, "PART[^LY]") ~ "PART",
    TRUE ~ NA_character_)) %>% 
  fill(chapter) %>% 
  filter(chapter != "PART") %>% 
  mutate(chapter = as.numeric(chapter))
```

```{r, warning = FALSE, message = FALSE}
#tokenize, remove gb_id
ti_tokens <- ti_df %>% 
  unnest_tokens(word, text) %>% 
  dplyr::select(-gutenberg_id)
```

```{r, warning = FALSE, message = FALSE}
#remove stop words
ti_nonstop_words <- ti_tokens %>% 
  anti_join(stop_words)

#word count sans stop words
ti_nonstop_counts <- ti_nonstop_words %>% 
  count(chapter, word)

#top 5 words per chapter
ti_top_5_words <- ti_nonstop_counts %>% 
  group_by(chapter) %>% 
  arrange(-n) %>% 
  slice(1:5)

```
### Top 5 Words by Chapter

This analysis looks at the top 5 words for each chapter of *Treasure Island* to associate any trends... The results were unsurprising, the book is about pirates after all.

```{r, fig.height= 10, fig.width= 10, warning = FALSE, message = FALSE, fig.cap = "**Figure 1:** Counts of the top 5 most frequently used words in Robert Louis Stevenson's *Treasure Island* by chapter."}
#visualize the word counts
ggplot(data = ti_top_5_words, aes(x = word, y = n)) +
  geom_col(fill = "blue") +
  facet_wrap(~chapter, scales = "free") +
  coord_flip() +
  theme_bw()
```

```{r, warning = FALSE, message = FALSE, results='hide'}
#join sentiment analysis lexicon to TI df
ti_afinn <- ti_nonstop_words %>% 
  inner_join(get_sentiments("afinn"))

#get afinn counts for chapter
afinn_counts <- ti_afinn %>% 
  count(chapter, value)
```

```{r, results='hide', warning=FALSE, message=FALSE}
# Plot them: 
#ggplot(data = afinn_counts, aes(x = value, y = n)) +
 # geom_col() +
  #facet_wrap(~chapter)
```
### Sentiment Analysis using AFINN

I ran the book against the AFINN lexicon to find the mean sentiment scores per chapter.  The book is a little on the dark side, especially for a book written primarily for boys.

```{r, message=FALSE, warning=FALSE, fig.width=6.5, fig.height=6.5, fig.cap= "**Figure 2:** Mean sentiment score for each chapter of Robert Louis Stevenson's *Treasure Island* based on the AFINN sentiment lexicon.  A mean score of -5 is a highly negative sentiment while a mean score of 5 is a highly positive sentiment"}
# Find the mean afinn score by chapter: 
afinn_means <- ti_afinn %>% 
  group_by(chapter) %>% 
  summarize(mean_afinn = mean(value))

#plot the afinn means
ggplot(data = afinn_means, 
       aes(x = fct_rev(as.factor(chapter)), y = mean_afinn)) +
  geom_col() +
  labs(y = "Mean AFINN integer", x = "Chapter") +
  coord_flip() +
  theme_bw()
```

### Word Cloud

The final product is a word cloud to present the top 100 words in the novel.

```{r, warning = FALSE, message = FALSE, preview = TRUE}
#create subset of top 100 words
ti_top100 <- ti_nonstop_words %>%
  count(word) %>% 
  arrange(-n) %>% 
  slice(1:100)

# Import image for background
img.file <- here("_posts/2021-03-14-treasure-island/", "treasure_island.jpg")
img <- readJPEG(img.file)

#plot it
ti_cloud <- ggplot(data = ti_top100, aes(label = word)) +
  background_image(img) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "circle") +
  scale_size_area(max_size = 10) +
  scale_color_gradientn(colors = c("goldenrod", "purple", "#660099")) +
  theme_minimal()

ti_cloud
```

### Citation

Stevenson, R. L. (2006) *Treasure Island.* Project Gutenberg. [Pdf] Retrieved February 20, 2021, from Project Gutenberg Mirror, http://mirrors.xmission.com/gutenberg/.
